---
title: "Global Terrorism Database (1970 - 2015) Preliminary Data Cleaning"
date: 2017-05-30
tags:
  - R-programming
  - Data Quality
categories:
  - Project
output: 
  md_document:
    variant: markdown_github
---
# Introduction

In the coming three posts, I will show my explrative data analytics work on the [Global Terrorism Database (GTD)](https://www.start.umd.edu/gtd/), a database maintained by the [National Consortium for the Study of Terrorism (START)](http://www.start.umd.edu/) at the University of Maryland, College Park. I used the [version published on Kaggle ](https://www.kaggle.com/START-UMD/gtd) in early 2017 for this project, which covers terrorist attacks happened between 1970 and 2015. (This database has been updated in July 2017 to include incidents in 2016.)

From the GTD exlporative analysis, I hoped to gather some insights from the 45 years of data on terrorism and global public safety. More specifically, I intend to answer the following questions:

* Since 1970, how many terrorist attacks have occurred in the United States? How does it look from a global scale?
* Of those attacks happened in the U.S., what are their attack methods and targets?
* How did these attacks change over time? Where did these attacks happen and what were their motives? 

I selected this dataset, and these questions in particular, because of their relevancy to current debates. I am interested in finding out where are the significant terrorist threats coming from and how have these threats changed over the years. The database relies on materials that are ***publicly available and unclassified***, covering terrorism incidents that happened between 1 January 1970 and 31 December 2015, with more recent incidents not published by START yet. As of now, the database includes over 2,600 terrorist attacks which have occurred in the United States. Additionally, it includes information about the target, perpetrator, and motivations. This makes the dataset ideal for my purposes.

In this post, I will step through the data cleaning and preprocessing techniques with R programming on the GTD database to make it ready for explorative data analysis and visualizations.

```{r, message=FALSE, warning=FALSE, include=FALSE}
# Load packages
library(dplyr)
library(stringr)
library(lubridate)
library(ggplot2)
library(ggmap)
library(purrr)
library(purrrlyr)
library(tidyverse)
library(reshape2)
library(caret) # For data partition
library(plotly) # For interactive mapping and visuals
library(leaflet)
library(cowplot)
library(ggthemes)
library(viridis)
library(knitr)
library(scales)
library(gridExtra)
library(MASS)
library(ggmap)
library(forcats)
library(tm)
library(wordcloud)
library(RColorBrewer)
```
## Load database into R environment
First, I will load the flat file from the original GTD database into an dataframe object.
```{r}
# Load data
db <- read.csv("globalterrorismdb_0616dist.csv", na.string = c("", "NA", " "))
class(db)
```
This dataframe object contains `r nrow(db)` incidents from the year of `r min(unique(db$iyear))` to the year of `r max(unique(db$iyear))`, with `r ncol(db)` variables. There are `r sum(db$country_txt == "United States")` incidents that happened in the U.S., while around `r sum(db[db$country_txt == "United States",]$doubtterr == 0)` cases are surely terrorism incidents. Since our core question is where the terrorism attacks in the US originate from, we will make these cases our primary focus on this project and trim down the incidents and variables accordingly.

## Clean up data type and format based on [GTD Code Book](https://www.start.umd.edu/gtd/downloads/Codebook.pdf)
```{r}
# Create a dataframe for attacks in the US
dt <- db %>%
  dplyr::select(eventid, iyear, imonth, iday, extended, summary, doubtterr, multiple, related, country_txt, region_txt, provstate, latitude, longitude, attacktype1_txt, attacktype2_txt, attacktype3_txt, success, suicide, weaptype1_txt, weaptype2_txt, weaptype3_txt, weaptype4_txt, target1, targtype1_txt, natlty1_txt, target2, targtype2_txt, natlty2_txt, target3, targtype3_txt,natlty3_txt, gname, gname2, gname3, guncertain1, guncertain2, guncertain3, nperps, nperpcap, claimed, compclaim, motive, nkill, nkillter, nwound, nwoundte, property, propextent, ishostkid, nhostkid, INT_LOG, INT_IDEO, INT_MISC, INT_ANY) %>% 
  filter(doubtterr == 0, country_txt == "United States")

# Create a dataframe with same structure for the world, for reference purpose
dt_w <- db %>%
  dplyr::select(eventid, iyear, imonth, iday, extended, summary, doubtterr, multiple, related, country_txt, region_txt, provstate, latitude, longitude, attacktype1_txt, attacktype2_txt, attacktype3_txt, success, suicide, weaptype1_txt, weaptype2_txt, weaptype3_txt, weaptype4_txt, target1, targtype1_txt, natlty1_txt, target2, targtype2_txt, natlty2_txt, target3, targtype3_txt,natlty3_txt, gname, gname2, gname3, guncertain1, guncertain2, guncertain3, nperps, nperpcap, claimed, compclaim, motive, nkill, nkillter, nwound, nwoundte, property, propextent, ishostkid, nhostkid, INT_LOG, INT_IDEO, INT_MISC, INT_ANY) %>% 
  filter(doubtterr == 0) # 217 is the country code for United States
```
The water-downed version of the database contains `r nrow(dt)` incidents and `r ncol(dt)` variables. 

### Check out the variables

I selected the most relevant details in the database for my EDA purposes. For each incident, the database includes variables that describe the incidents from nine different aspects:

Incident Aspect | Variable Names | Incident Aspect | Variable Names
----------------|----------------|-----------------|---------------
GTD ID and Date | eventid, iyear, imonth, iday, extended | Target/Victim Information | target1, targtype1_txt, natlty1_txt, target2, targtype2_txt, natlty2_txt, target3, targtype3_txt, natlty3_txt
Incident Information | summary, doubtterr, multiple, related | Perpetrator Information | gname, gname2, gname3, guncertain1, guncertain2, guncertain3, nperps, nperpcap, claimed, compclaim
Incident Location | country_txt, region_txt, provstate, latitude, longitude | Casualties and Consequences |  nkill, nkillter, nwound, nwoundte, property, propextent, ishostkid, nhostkid
Attack Information | attacktype1_txt, attacktype2_txt, attacktype3_txt, success, suicide | Additional Information and Sources | INT_LOG, INT_IDEO, INT_MISC, INT_ANY
Weapon Information | weaptype1_txt, weaptype2_txt, weaptype3_txt, weaptype4_txt

Please refer to the GTD code book if you are interested to know more about each variable listed here.

### Double check variable types and deal with data quality

My next step is to make sure the data I am using are of consistent format. I will also create a few extra variables to fulfill following steps in data exploration.

After a quick glimpse at the data, I should conduct the following changes to the dataframe:
```{r}
# Change variables with strings back to character vector
dt$summary <- as.character(dt$summary)
dt$target1 <- as.character(dt$target1)
dt$target2 <- as.character(dt$target2)
dt$target3 <- as.character(dt$target3)

# Trim factor variables to only have included levels
dt <- dt %>% dmap_if(is.factor, fct_drop)

# Recode -9 and -99 as NA in the dataframe
dt[dt == -9 | dt == -99] <- NA

# Recode factors "." and "Unknown" into NA in the factors
for (i in 1:ncol(dt)){
  if (is.factor(dt[,i])){
     levels(dt[,i]) <- sub("^.$", NA, levels(dt[,i]))
     levels(dt[,i]) <- sub("Unknown", NA, levels(dt[,i]))
  }
}
```

```{r}
# Run the same codes on the world reference dataframe
# Change variables with strings back to character vector
dt_w$summary <- as.character(dt_w$summary)
dt_w$target1 <- as.character(dt_w$target1)
dt_w$target2 <- as.character(dt_w$target2)
dt_w$target3 <- as.character(dt_w$target3)
# Trim factor variables to only have included levels
dt_w <- dt_w %>% dmap_if(is.factor, fct_drop)
# Recode -9 and -99 as NA in the dataframe
dt_w[dt_w == -9 | dt_w == -99] <- NA
# Recode factors "." and "Unknown" into NA in the factors
for (i in 1:ncol(dt_w)){
  if (is.factor(dt_w[,i])){
     levels(dt_w[,i]) <- sub("^.$", NA, levels(dt_w[,i]))
     levels(dt_w[,i]) <- sub("Unknown", NA, levels(dt_w[,i]))
  }
}

```
Now that the data is cleaned and unified, I will generate a **date** feature that combines the *iyear, imonth, and iday features* in the dateframe for easier usage in date slicing. Since the dataframe has `r sum(dt$iday == 0)` incidents that does not have a clarified day, I will assign 1 to these incidents as an approximation in order to create a complete date object. (This won't be a problem for my data exploration since I will only be using year and month on such a wide time span.)

```{r}
# Replace unknown days (0 according to the code book) with 1
dt$iday <- as.integer(gsub(0, 1, dt$iday))

# Create a new variable "idate"
dt$idate <- as.Date(paste0(dt$iyear,
                       stringr::str_pad(as.character(dt$imonth), width = 2, side = "left", pad = "0"),
                       stringr::str_pad(as.character(dt$iday), width = 2, side = "left", pad = "0")), "%Y%m%d")
# There are 23 incidents where the date object fails to show
# a closer look finds that these are incidents where iday = 31
# when the month actually won't have 31 days.
# We will treat these idays as 30, i.e. end of the month
dt$iday[is.na(dt$idate)] <- 30
dt$idate <- as.Date(paste0(dt$iyear,
                       stringr::str_pad(as.character(dt$imonth), width = 2, side = "left", pad = "0"),
                       stringr::str_pad(as.character(dt$iday), width = 2, side = "left", pad = "0")), "%Y%m%d")
```

```{r}
# Repeat Same Step for the World Refernce Dataframe
# Replace unknown days (0 according to the code book) with 1
dt_w$iday <- as.integer(gsub(0, 1, dt_w$iday))

# Create a new variable "idate"
dt_w$idate <- as.Date(paste0(dt_w$iyear,
                       stringr::str_pad(as.character(dt_w$imonth), width = 2, side = "left", pad = "0"),
                       stringr::str_pad(as.character(dt_w$iday), width = 2, side = "left", pad = "0")), "%Y%m%d")
# There are 23 incidents where the date object fails to show
# a closer look finds that these are incidents where iday = 31
# when the month actually won't have 31 days.
# We will treat these idays as 30, i.e. end of the month
dt_w$iday[is.na(dt_w$idate)] <- 30
dt_w$idate <- as.Date(paste0(dt_w$iyear,
                       stringr::str_pad(as.character(dt_w$imonth), width = 2, side = "left", pad = "0"),
                       stringr::str_pad(as.character(dt_w$iday), width = 2, side = "left", pad = "0")), "%Y%m%d")
```

The polished dataframe should have variables that look like below:

```{r}
#Take a quick look at the variables
glimpse(dt)
```
